Complete draft

### 2019-05-09 doc/point.tex
update but need to check Fisher matrix

There is sufficient information in the distribution of peculiar velocities and overdensities of galaxies to test general relativity on the largest scale. Suppose we have a region that encloses $N_{gal}$ number of galaxies in the universe. We can divide the region into $N_{cell}\gg 1$ number of equal volume cells whose size can be made so small that either one or no galaxy is contained in one cell. This is encoded in a vector $\textbf{N}$ of size $N_{cell}$ whose entries are either 0 or 1. For any 2 cells, we can compute their correlation, a function of the distance $r$ between the cells only. In our application, we consider $\xi_{\delta\delta}$ (the overdensity-overdensity correlation), $\xi_{\delta v}$ (overdensity-velocity correlation), $\xi_{v\delta}$ (velocity-overdensity correlation), and $\xi_{vv}$ (velocity-velocity correlation).\par
We postulate that deviations from expected variance $\bar{N}$ and expected peculiar velocity $\bar{v}$ are Gaussian. Then the above correlation matrices can be used to construct a block matrix $\textbf{C}$, parameterized by $A$ and $\bm{\sigma}_v$, a vector of shot noise for peculiar velocity measurement:
\begin{align}
\textbf{C} &\equiv \textbf{K} + A\textbf{S} \\ 
&=
\left(
\begin{array}{cc}
\bar{N}\textbf{I} & 0 \nonumber \\
0 & \bm{\sigma}_v^2\textbf{I}
\end{array}
\right)
+
A
\left(
\begin{array}{cc}
\bar{N}^2\xi_{\delta\delta} & \bar{N}\xi_{\delta v} \\
\bar{N}\xi_{\delta v} & \xi_{vv}
\end{array}
\right).
\end{align}\par
The reason we have $\bar{N}$ as the variance for overdensity is that our small galaxy cells are best modeled by Poisson statistics, in which the mean equals the variance. In addition, note that the dimension of $C$ is $N_{cell}+N_{gal}$ by $N_{cell}+N_{gal}$. For the peculiar velocity covariance, we assign $\sigma^2=\infty$ to cells that do not contain any galaxy, and one can show that removing rows and columns that contain those cells has no effect but a constant offset on what we will compute later. \par
Given the block matrix $C$ and our data $\textbf{x}= [N_1-\bar{N}, N_2-\bar{N}, \hdots, N_{N_{cell}}-\bar{N}, \Delta v_1, \hdots, \Delta v_{N_{gal}}]^T$, where $\Delta v$ is the deviation of the measured velocity from the expected, we can form the Gaussian likelihood as a function of $A$
\[
\mathcal{L}(A) = \frac{1}{\sqrt{\left(2\pi\right)^{N_{cell}+N_{gal}} \det \textbf{C}}}\exp\left\{-\frac{1}{2}\textbf{x}^T\textbf{C}^{-1}\textbf{x}\right\}.
\]\par
To make the expression cleaner, we take the log of the likelihood and focus on the $A$-dependent part:
\begin{equation}
-2\ln\mathcal{L}(A) = \ln\det\textbf{C} + \textbf{x}^T\textbf{C}^{-1}\textbf{x} + \text{const}.
\label{log:eqn}
\end{equation}\par

%  ---- Change
%NEW STUFF

We new expand the log-likelihood as a Taylor series for small $A$ around $A=0$.

First, we focus on the  term $\textbf{x}^T\textbf{C}^{-1}\textbf{x}$. We rewrite $\textbf{C}$ by normalizing out its uncorrelated $\textbf{K}$ term
by   introducing the matrices $\textbf{M}$ and $\textbf{X}$,
where $\textbf{M}=\sqrt{\textbf{K}}$ and $\textbf{S} = \textbf{M} \textbf{X} \textbf{M}$ such that $\textbf{X}=\textbf{M}^{-1}\textbf{S}\textbf{M}^{-1}$. Now, our matrix $\textbf{C}$ can be re-written in terms of $\textbf{X}$ and $\textbf{M}$:
\begin{equation}
\textbf{C} = \textbf{M}(\textbf{I}+A\textbf{X})\textbf{M}.
\end{equation}
Expanded as a Taylor series
\begin{align}
\textbf{C}^{-1} & = \textbf{M}^{-1}(1-A\textbf{X}+A^2 \textbf{X}^2 + \ldots) \textbf{M}^{-1} \\
& = \textbf{K}^{-1}   -A \textbf{K}^{-1}\textbf{S}\textbf{K}^{-1} +\frac{A^2}{2} \textbf{K}^{-1}\textbf{S}\textbf{K}^{-1}\textbf{S}\textbf{K}^{-1} +O(A^3) + \ldots 
\end{align}
%= -A \textbf{K}^{-1}\textbf{S}\textbf{K}^{-1} + \text{const}.
%\] \par

We evaluate $\textbf{x}^T \textbf{C}^{-1}\textbf{x}$ by splitting up the counts and velocities.  The data vector $\textbf{x}$ is composed of the counts
 $\textbf{N}-\bar{N}$ and peculiar velocities $\mathbf{\Delta v}$, whereas $\textbf{S}$ is comprised of $\bar{N}^2\xi_{\delta\delta}$, $\bar{N}\xi_{\delta v}$,
and $\xi_{vv}$.  In this manner
\begin{align}
\textbf{x}^T \textbf{K}^{-1}\textbf{S}\textbf{K}^{-1} \textbf{x}& = (\textbf{N}-\bar{N}) \xi_{\delta\delta}  (\textbf{N}-\bar{N})  + 2  (\textbf{N}-\bar{N}) \xi_{\delta v}\frac{\mathbf{\Delta v}}{\bm{\sigma}_v^2}+ 
\frac{\mathbf{\Delta v}}{\bm{\sigma}_v^2} \xi_{vv}\frac{\mathbf{\Delta v}}{\bm{\sigma}_v^2}.
\end{align}
Noting that $N_i=1$ when the cell $i$ contains a galaxy and zero otherwise and that only cells with a peculiar velocity measurement contribute ($\sigma_v =\infty$
for cells without such a measurement), and that in the limit of infinitely small cells, $\bar{N}\sum_{i,j\in N_{cell}} = n \int dV$, where $n$ is the galaxy number density,
\begin{align*}
\textbf{x}^T \textbf{K}^{-1}\textbf{S}\textbf{K}^{-1} \textbf{x}& = \sum_{i,j\in N_{gal}} \xi_{\delta\delta,ij} -2 \bar{N} \sum_{i\in N_{gal},j\in N_{cell}} \xi_{\delta\delta,ij}\\
& +\bar{N}^2  \sum_{i,j\in N_{cell}}  \xi_{\delta\delta,ij} + 2 \sum_{i,j\in N_{gal}} \frac{\Delta v_j}{\sigma_{v,j}^2} \xi_{\delta v,ij} \\
& -2\bar{N}\sum_{i\in N_{cell}, j\in N_{gal}} \frac{\Delta v_j}{\sigma_{v,j}^2} \xi_{\delta v,ij}  + \sum_{i,j\in N_{gal}} \frac{\Delta v_i}{\sigma_{v,i}^2} \xi_{vv,ij} \frac{\Delta v_j}{\sigma_{v,j}^2}\\
& = \sum_{i,j\in N_{gal}} \left( \xi_{\delta\delta,ij} + 2  \frac{\Delta v_j}{\sigma_{v,j}^2} \xi_{\delta v,ij} + \frac{\Delta v_i}{\sigma_{v,i}^2} \xi_{vv,ij} \frac{\Delta v_j}{\sigma_{v,j}^2}\right)\\
&-2 n \sum_{i\in N_{gal}}\int \left(\xi_{\delta\delta,iV} + \frac{\Delta v_i}{\sigma_{v,i}^2} \xi_{v \delta,iV}\right) dV\\
& +n^2  \int \xi_{\delta\delta,V_1V_2}dV_1dV_2.
\end{align*}

Now consider the next term $\textbf{K}^{-1}\textbf{S}\textbf{K}^{-1}\textbf{S}\textbf{K}^{-1}$.  The middle section is
\begin{align}
\textbf{S}\textbf{K}^{-1}\textbf{S} & =
\begin{pmatrix}
 \bar{N}^2  \xi_{\delta v}\frac{1}{\bm{\sigma}_v^2}  \xi_{v \delta}  & \bar{N} \xi_{\delta v} \frac{1} {\bm{\sigma}_v^2} \xi_{vv}  \\
 \bar{N} \xi_{v v} \frac{1} {\bm{\sigma}_v^2} \xi_{v \delta}   &  \xi_{vv} \frac{1}{\bm{\sigma}_v^2} \xi_{vv}
\end{pmatrix}
\end{align}
to leading order of $\bar{N}$ anticipating that we will take the limit where it goes to zero.
Then
\begin{align}
\textbf{x}^T \textbf{K}^{-1}\textbf{S}\textbf{K}^{-1}\textbf{S}\textbf{K}^{-1} \textbf{x}
& = \sum_{i,j\in N_{gal}} \left( \left(\xi_{\delta v}\frac{1}{\bm{\sigma}_v^2}  \xi_{v \delta}\right)_{ij}
 + 2  \frac{\Delta v_j}{\sigma_{v,j}^2} \left(\xi_{\delta v} \frac{1} {\bm{\sigma}_v^2} \xi_{vv}  \right)_{ij} + \frac{\Delta v_i}{\sigma_{v,i}^2} \left( \xi_{vv} \frac{1}{\bm{\sigma}_v^2} \xi_{vv}\right)_{ij} \frac{\Delta v_j}{\sigma_{v,j}^2}\right)\\
&-2 n \sum_{i\in N_{gal}}\int \left( \left(\xi_{\delta v}\frac{1}{\bm{\sigma}_v^2}  \xi_{v \delta}\right)_{iV} + \frac{\Delta v_i}{\sigma_{v,i}^2} \left( \xi_{v v} \frac{1} {\bm{\sigma}_v^2} \xi_{v \delta}  \right)_{iV} \right) dV\\
& +n^2  \int  \left(\xi_{\delta v}\frac{1}{\bm{\sigma}_v^2}  \xi_{v \delta}\right)_{V_1V_2}dV_1dV_2.
\end{align}

@@

\begin{comment}
END NEW STUFF

@@

First, we focus on the second term, $\textbf{x}^T\textbf{C}^{-1}\textbf{x}$. Looking at the definition of $\textbf{C}$ in equation (1), we want to factor $\textbf{K}$ into an identity matrix. To do so, we introduce matrix $\textbf{M}$ and $\textbf{X}$. We let $\textbf{M}=\sqrt{\textbf{K}}$, and $\textbf{S} = \textbf{M} \textbf{X} \textbf{M}$. Then $\textbf{X}=\textbf{M}^{-1}\textbf{S}\textbf{M}^{-1}$. Now, our matrix $\textbf{C}$ can be re-written in terms of $\textbf{X}$ and $\textbf{M}$:

@@

\textbf{C}^{-1} = \textbf{M}^{-1}(1-A\textbf{X}) \textbf{M}^{-1} = -A \textbf{K}^{-1}\textbf{S}\textbf{K}^{-1} + \text{const}.

@@

&= -A\left(\textbf{x}^T \textbf{K}^{-1}\right)\textbf{S}\left(\textbf{K}^{-1}\textbf{x}\right).

@@

-\left(\textbf{x}^T \textbf{K}^{-1}\right)\textbf{S}\left(\textbf{K}^{-1}\textbf{x}\right) = s_1 +s_2+\hdots +s_6

@@

&=\textbf{K}^{-1}\textbf{S}\textbf{K}^{-1}\textbf{S}\textbf{K}^{-1}

@@

I first sight I wanted to move $\textbf{S}$ across the $\textbf{K}^{-1}$ by using a commutator, thinking that it might simplify things a lot. But it turned out that the commutator did not lead to the expected cancelation, so let's do it the hard way. We first do $\textbf{K}^{-1}\textbf{S}\textbf{K}^{-1}\textbf{S}$, henceforth denoting the inverse of $\bm{\sigma}^2_v\textbf{I}$ as $\bm{\sigma}^{-1}$:

@@

\textbf{K}^{-1}\textbf{S}\textbf{K}^{-1}\textbf{S} &=

@@

Next, we multiply $\textbf{K}^{-1}$ from the right and get

@@

\textbf{K}^{-1}\textbf{S}\textbf{K}^{-1}\textbf{S}\textbf{K}^{-1} &=

@@

\textbf{K}^{-1}\textbf{S}\textbf{K}^{-1}\textbf{S}\textbf{K}^{-1}=

@@

Having obtained the 4 block matrices, we now proceed with the calculation of $\textbf{x}^T\textbf{K}^{-1}\textbf{S}\textbf{K}^{-1}\textbf{S}\textbf{K}^{-1}\textbf{x}$. Similar to what we did for the first order, we break the sum into the following terms (again, summation is within the range of $N_{gal}$ by default unless otherwise stated):

@@

&= -\frac{1}{2}\tr(\textbf{K}^{-1}\textbf{S}\textbf{K}^{-1}\textbf{S}) \\

>>>>

There is sufficient information in the distribution of peculiar velocities and overdensities of galaxies to test general relativity on the largest scale. Suppose we have a region that encloses $N_{gal}$ number of galaxies in the universe. We can divide the region into $N_{cell}\gg 1$ number of equal volume cells whose size can be made so small that either one or no galaxy is contained in one cell. This is encoded in a vector $\textbf{N}$ of size $N_{cell}$ whose entries are either 0 or 1. For any 2 cells, we can compute their correlation, a function of the distance $r$ between the cells only. In our application, we consider $\xi_{\delta\delta}$ (the overdensity-overdensity correlation), $\xi_{\delta v}$ (overdensity-velocity correlation), $\xi_{v\delta}$ (velocity-overdensity correlation), and $\xi_{vv}$ (velocity-velocity correlation).\par

The data  is $\textbf{x}= [N_1-\bar{N}, N_2-\bar{N}, \hdots, N_{N_{cell}}-\bar{N}, \Delta v_1, \hdots, \Delta v_{N_{gal}}]^T$, where $\Delta v$ is the deviation of the measured velocity from the expected,

We postulate that deviations from expected variance $\bar{N}$ and expected peculiar velocity $\bar{v}$ are Gaussian. Then the above correlation matrices can be used to construct a block matrix $\textbf{C}$, parameterized by $A$ and $\bm{\sigma}_v$, a vector of shot noise for peculiar velocity measurement:
\begin{align}
\textbf{C} &\equiv \textbf{C}_0 + \textbf{S}\\
%\textbf{C} &\equiv \textbf{C}_0 + A\textbf{S} \\ 
&=
\left(
\begin{array}{cc}
\bar{N}\textbf{I} & 0 \nonumber \\
0 & \bm{\sigma}_v^2\textbf{I}
\end{array}
\right)
+
\begin{pmatrix}
\bar{N}^2\xi_{gg} & \bar{N}\xi_{g v} \\
\bar{N}\xi_{v g} & \xi_{vv}
\end{pmatrix},
\end{align}
where $\xi_{gg}$, $\xi_{vv}$, and $\bar{N}\xi_{g v}$ are the galaxy overdensity and peculiar velocity correlation functions and their cross-correlation
respectively.

The reason we have $\bar{N}$ as the variance for overdensity is that our small galaxy cells are best modeled by Poisson statistics, in which the mean equals the variance. In addition, note that the dimension of $C$ is $N_{cell}+N_{gal}$ by $N_{cell}+N_{gal}$. For the peculiar velocity covariance, we assign $\sigma^2=\infty$ to cells that do not contain any galaxy, and one can show that removing rows and columns that contain those cells has no effect but a constant offset on what we will compute later.

The science parameters $\theta$ enter through these correlation functions, and may include $\theta \in \{fD, bD, \gamma, \Omega_{M_0},\ldots\}$ and others
that describe the matter power spectrum.
For example, the parameter $fD$ enters in through the relations $\xi_{vv}\propto (fD\mu)^2$, the SN~Ia host-galaxy count overdensity
power spectrum $\xi_{g g }\propto (bD + fD\mu^2)^2$, and the galaxy-velocity cross-correlation $\xi_{vg}
\propto  (bD + fD\mu^2)fD$, where $b$ is the galaxy bias and $\mu\equiv \cos{(\hat{k} \cdot \hat{r})}$ where $\hat{r}$ is the direction of
the line of sight.  
Assuming linear dependence of the parameters,
\begin{align}
\textbf{C} 
&=
\left(
\begin{array}{cc}
\bar{N}\textbf{I} & 0 \nonumber \\
0 & \bm{\sigma}_v^2\textbf{I}
\end{array}
\right)
+ \sum_\theta \theta
\begin{pmatrix}
\bar{N}^2\xi_{gg,\theta} & \bar{N}\xi_{g v,\theta} \\
\bar{N}\xi_{v g,\theta} & \xi_{vv,\theta}
\end{pmatrix}\\
&=
\left(
\begin{array}{cc}
\bar{N}\textbf{I} & 0 \nonumber \\
0 & \bm{\sigma}_v^2\textbf{I}
\end{array}
\right)
+ \sum_\alpha \theta_\alpha \textbf{S}_\alpha .
\end{align}

%\left(
%\begin{array}{cc}
%\bar{N}\textbf{I} & 0 \nonumber \\
%0 & \bm{\sigma}_v^2\textbf{I}
%\end{array}
%\right)
%+
%A
%\left(
%\begin{array}{cc}
%\bar{N}^2\xi_{\delta\delta} & \bar{N}\xi_{\delta v} \\
%\bar{N}\xi_{\delta v} & \xi_{vv}
%\end{array}
%\right).
%\end{align}\par

The Gaussian likelihood as a function of $\theta$
\[
\mathcal{L}(\theta) = \frac{1}{\sqrt{\left(2\pi\right)^{N_{cell}+N_{gal}} \det \textbf{C}}}\exp\left\{-\frac{1}{2}\textbf{x}^T\textbf{C}^{-1}\textbf{x}\right\}.
\]\par
To make the expression cleaner, we take the log of the likelihood and focus on the $A$-dependent part:
\begin{equation}
-2\ln\mathcal{L}(\theta) = \ln\det\textbf{C} + \textbf{x}^T\textbf{C}^{-1}\textbf{x} + \text{const}.
\label{log:eqn}
\end{equation}

The estimator and Fisher Matrix for this Gaussian likelihood are known
\begin{equation}
\hat{\theta}_\alpha = \frac{1}{2} F^{-1}_{\alpha \beta} [(\textbf{x}^T \textbf{C}^{-1} \textbf{S}_\beta \textbf{C}^{-1} \textbf{x}) - \tr(\textbf{C}^{-1} \textbf{S}_\beta \textbf{C}^{-1} \textbf{C}_0)]
\end{equation}
and
\begin{equation}
F_{\alpha\beta} = \frac{1}{2} \tr(\textbf{C}^{-1} \textbf{S}_\beta \textbf{C}^{-1} \textbf{C}_0 \textbf{S}_\beta).
\end{equation}

The term
\begin{align}
\textbf{x}^T \textbf{C}^{-1} \textbf{S}_\beta \textbf{C}^{-1} \textbf{x} & \approx \textbf{x}^T \textbf{C}_0^{-1} \textbf{S}_\beta \textbf{C}_0^{-1} \textbf{x}\\
&=(\textbf{N}-\bar{N}) \xi_{\delta\delta,\theta_\beta}  (\textbf{N}-\bar{N})  + 2  (\textbf{N}-\bar{N}) \xi_{\delta v,\theta_\beta}\frac{\mathbf{\Delta v}}{\bm{\sigma}_v^2}+ 
\frac{\mathbf{\Delta v}}{\bm{\sigma}_v^2} \xi_{vv,\theta_\beta}\frac{\mathbf{\Delta v}}{\bm{\sigma}_v^2}.
\end{align}
Noting that $N_i=1$ when the cell $i$ contains a galaxy and zero otherwise and that only cells with a peculiar velocity measurement contribute ($\sigma_v =\infty$
for cells without such a measurement), and that in the limit of infinitely small cells, $\bar{N}\sum_{i,j\in N_{cell}} = \int n dV$, where $n$ is the galaxy number density,
\begin{align*}
\textbf{x}^T \textbf{C}^{-1} \textbf{S}_\beta \textbf{C}^{-1} \textbf{x} & \approx \sum_{i,j\in N_{gal}} \xi_{\delta\delta,\theta_\beta,ij} -2 \bar{N} \sum_{i\in N_{gal},j\in N_{cell}} \xi_{\delta\delta,\theta_\beta,ij}\\
& +\bar{N}^2  \sum_{i,j\in N_{cell}}  \xi_{\delta\delta,\theta_\beta,ij} + 2 \sum_{i,j\in N_{gal}} \frac{\Delta v_j}{\sigma_{v,j}^2} \xi_{\delta v,\theta_\beta,ij} \\
& -2\bar{N}\sum_{i\in N_{cell}, j\in N_{gal}} \frac{\Delta v_j}{\sigma_{v,j}^2} \xi_{\delta v,\theta_\beta,ij}  + \sum_{i,j\in N_{gal}} \frac{\Delta v_i}{\sigma_{v,i}^2} \xi_{vv,\theta_\beta,ij} \frac{\Delta v_j}{\sigma_{v,j}^2}\\
& = \sum_{i,j\in N_{gal}} \left( \xi_{\delta\delta,\theta_\beta}(x_i,x_j) + 2  \frac{\Delta v_j}{\sigma_{v,j}^2} \xi_{\delta v,\theta_\beta}(x_i,x_j) + \frac{\Delta v_i}{\sigma_{v,i}^2} \xi_{vv,\theta_\beta}(x_i,x_j) \frac{\Delta v_j}{\sigma_{v,j}^2}\right)\\
&-2  \sum_{i\in N_{gal}}\int n(x_i) \left(\xi_{\delta\delta,\theta_\beta}(x_i,y) + \frac{\Delta v_i}{\sigma_{v,i}^2} \xi_{v \delta,\theta_\beta}(x_i,y) \right) d^3y\\
&  + \int n(x) n(y) \xi_{\delta\delta,\theta_\beta}(x,y)d^3xd^3y.
\end{align*}

The trace term is
\begin{align}
 \tr(\textbf{C}^{-1} \textbf{S}_\beta \textbf{C}^{-1} \textbf{C}_0) & \approx  \tr(\textbf{C}_0^{-1} \textbf{S}_\beta) \\
 & = \bar{N} \tr(\xi_{\delta\delta,\beta}) + \tr(\frac{1}{\sigma^2_v} \xi_{vv,\beta})\\
 & = \int n(x) \xi_{\delta \delta,\beta}(x,x) d^3x +  \sum_{i\in N_{gal}} \frac{1}{\sigma^2_{v,i}} \xi_{vv,\beta}(x_i,x_i)
\end{align}

The remaining term to the leading term in $\bar{N}$ is
\begin{align}
 \tr(\textbf{C}^{-1} \textbf{S}_\alpha \textbf{C}^{-1} \textbf{S}_\beta) & \approx  \tr(\textbf{C}_0^{-1} \textbf{S}_\alpha \textbf{C}_0^{-1} \textbf{S}_\beta)  \\
 & = \bar{N} \tr(\xi_{\delta v,\alpha} \frac{1}{\bm{\sigma}^2_v} \xi_{v\delta,\beta} ) + \tr(\frac{1}{\bm{\sigma}^2_v} \xi_{vv,\alpha}  \frac{1}{\bm{\sigma}^2_v}  \xi_{vv,\beta})\\
 & = \int n(x) \sum_{i\in N_{gal}} \xi_{\delta v,\alpha}(x,x_i) \frac{1}{\sigma^2_{v,i}} \xi_{v\delta,\beta} (x_i,x) d^3x \\
 & +  \sum_{i,j\in N_{gal}} \frac{1}{\sigma^2_{v,i}} \xi_{vv,\alpha}(x_i,x_j)  \frac{1}{\sigma^2_{v,j}}  \xi_{vv,\beta}(x_j,x_i).
\end{align}

%  ---- Change
%NEW STUFF
\begin{comment}
We now expand the log-likelihood as a Taylor series for small $\theta$ around $\theta=0$.

First, we focus on the  term $\textbf{x}^T\textbf{C}^{-1}\textbf{x}$. We rewrite $\textbf{C}$ by normalizing out its uncorrelated $\textbf{C}_0$ term
by   introducing the matrices $\textbf{M}$ and $\textbf{X}$,
where $\textbf{M}=\sqrt{\textbf{C}_0}$ and $\textbf{S} = \textbf{M} \textbf{X} \textbf{M}$ such that $\textbf{X}=\textbf{M}^{-1}\textbf{S}\textbf{M}^{-1}$. Now, our matrix $\textbf{C}$ can be re-written in terms of $\textbf{X}$ and $\textbf{M}$:
\begin{equation}
\textbf{C} = \textbf{M}(\textbf{I}+A\textbf{X})\textbf{M}.
\end{equation}
Expanded as a Taylor series
\begin{align}
\textbf{C}^{-1} & = \textbf{M}^{-1}(1-A\textbf{X}+A^2 \textbf{X}^2 + \ldots) \textbf{M}^{-1} \\
& = \textbf{C}_0^{-1}   -A \textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1} +\frac{A^2}{2} \textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1} +O(A^3) + \ldots 
\end{align}
%= -A \textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1} + \text{const}.
%\] \par

A term of interest is 
\begin{align}
\textbf{x}^T \textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1} \textbf{x}& = (\textbf{N}-\bar{N}) \xi_{\delta\delta}  (\textbf{N}-\bar{N})  + 2  (\textbf{N}-\bar{N}) \xi_{\delta v}\frac{\mathbf{\Delta v}}{\bm{\sigma}_v^2}+ 
\frac{\mathbf{\Delta v}}{\bm{\sigma}_v^2} \xi_{vv}\frac{\mathbf{\Delta v}}{\bm{\sigma}_v^2}.
\end{align}
Noting that $N_i=1$ when the cell $i$ contains a galaxy and zero otherwise and that only cells with a peculiar velocity measurement contribute ($\sigma_v =\infty$
for cells without such a measurement), and that in the limit of infinitely small cells, $\bar{N}\sum_{i,j\in N_{cell}} = n \int dV$, where $n$ is the galaxy number density,
\begin{align*}
\textbf{x}^T \textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1} \textbf{x}& = \sum_{i,j\in N_{gal}} \xi_{\delta\delta,ij} -2 \bar{N} \sum_{i\in N_{gal},j\in N_{cell}} \xi_{\delta\delta,ij}\\
& +\bar{N}^2  \sum_{i,j\in N_{cell}}  \xi_{\delta\delta,ij} + 2 \sum_{i,j\in N_{gal}} \frac{\Delta v_j}{\sigma_{v,j}^2} \xi_{\delta v,ij} \\
& -2\bar{N}\sum_{i\in N_{cell}, j\in N_{gal}} \frac{\Delta v_j}{\sigma_{v,j}^2} \xi_{\delta v,ij}  + \sum_{i,j\in N_{gal}} \frac{\Delta v_i}{\sigma_{v,i}^2} \xi_{vv,ij} \frac{\Delta v_j}{\sigma_{v,j}^2}\\
& = \sum_{i,j\in N_{gal}} \left( \xi_{\delta\delta,ij} + 2  \frac{\Delta v_j}{\sigma_{v,j}^2} \xi_{\delta v,ij} + \frac{\Delta v_i}{\sigma_{v,i}^2} \xi_{vv,ij} \frac{\Delta v_j}{\sigma_{v,j}^2}\right)\\
&-2 n \sum_{i\in N_{gal}}\int \left(\xi_{\delta\delta,iV} + \frac{\Delta v_i}{\sigma_{v,i}^2} \xi_{v \delta,iV}\right) dV\\
& +n^2  \int \xi_{\delta\delta,V_1V_2}dV_1dV_2.
\end{align*}

We evaluate $\textbf{x}^T \textbf{C}^{-1}\textbf{x}$ by splitting up the counts and velocities.  The data vector $\textbf{x}$ is composed of the counts
 $\textbf{N}-\bar{N}$ and peculiar velocities $\mathbf{\Delta v}$, whereas $\textbf{S}$ is comprised of $\bar{N}^2\xi_{\delta\delta}$, $\bar{N}\xi_{\delta v}$,
and $\xi_{vv}$.  In this manner
\begin{align}
\textbf{x}^T \textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1} \textbf{x}& = (\textbf{N}-\bar{N}) \xi_{\delta\delta}  (\textbf{N}-\bar{N})  + 2  (\textbf{N}-\bar{N}) \xi_{\delta v}\frac{\mathbf{\Delta v}}{\bm{\sigma}_v^2}+ 
\frac{\mathbf{\Delta v}}{\bm{\sigma}_v^2} \xi_{vv}\frac{\mathbf{\Delta v}}{\bm{\sigma}_v^2}.
\end{align}
Noting that $N_i=1$ when the cell $i$ contains a galaxy and zero otherwise and that only cells with a peculiar velocity measurement contribute ($\sigma_v =\infty$
for cells without such a measurement), and that in the limit of infinitely small cells, $\bar{N}\sum_{i,j\in N_{cell}} = n \int dV$, where $n$ is the galaxy number density,
\begin{align*}
\textbf{x}^T \textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1} \textbf{x}& = \sum_{i,j\in N_{gal}} \xi_{\delta\delta,ij} -2 \bar{N} \sum_{i\in N_{gal},j\in N_{cell}} \xi_{\delta\delta,ij}\\
& +\bar{N}^2  \sum_{i,j\in N_{cell}}  \xi_{\delta\delta,ij} + 2 \sum_{i,j\in N_{gal}} \frac{\Delta v_j}{\sigma_{v,j}^2} \xi_{\delta v,ij} \\
& -2\bar{N}\sum_{i\in N_{cell}, j\in N_{gal}} \frac{\Delta v_j}{\sigma_{v,j}^2} \xi_{\delta v,ij}  + \sum_{i,j\in N_{gal}} \frac{\Delta v_i}{\sigma_{v,i}^2} \xi_{vv,ij} \frac{\Delta v_j}{\sigma_{v,j}^2}\\
& = \sum_{i,j\in N_{gal}} \left( \xi_{\delta\delta,ij} + 2  \frac{\Delta v_j}{\sigma_{v,j}^2} \xi_{\delta v,ij} + \frac{\Delta v_i}{\sigma_{v,i}^2} \xi_{vv,ij} \frac{\Delta v_j}{\sigma_{v,j}^2}\right)\\
&-2 n \sum_{i\in N_{gal}}\int \left(\xi_{\delta\delta,iV} + \frac{\Delta v_i}{\sigma_{v,i}^2} \xi_{v \delta,iV}\right) dV\\
& +n^2  \int \xi_{\delta\delta,V_1V_2}dV_1dV_2.
\end{align*}

Now consider the next term $\textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1}$.  The middle section is
\begin{align}
\textbf{S}\textbf{C}_0^{-1}\textbf{S} & =
\begin{pmatrix}
 \bar{N}^2  \xi_{\delta v}\frac{1}{\bm{\sigma}_v^2}  \xi_{v \delta}  & \bar{N} \xi_{\delta v} \frac{1} {\bm{\sigma}_v^2} \xi_{vv}  \\
 \bar{N} \xi_{v v} \frac{1} {\bm{\sigma}_v^2} \xi_{v \delta}   &  \xi_{vv} \frac{1}{\bm{\sigma}_v^2} \xi_{vv}
\end{pmatrix}
\end{align}
to leading order of $\bar{N}$ anticipating that we will take the limit where it goes to zero.
Then
\begin{align}
\textbf{x}^T \textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1} \textbf{x}
& = \sum_{i,j\in N_{gal}} \left( \left(\xi_{\delta v}\frac{1}{\bm{\sigma}_v^2}  \xi_{v \delta}\right)_{ij}
 + 2  \frac{\Delta v_j}{\sigma_{v,j}^2} \left(\xi_{\delta v} \frac{1} {\bm{\sigma}_v^2} \xi_{vv}  \right)_{ij} + \frac{\Delta v_i}{\sigma_{v,i}^2} \left( \xi_{vv} \frac{1}{\bm{\sigma}_v^2} \xi_{vv}\right)_{ij} \frac{\Delta v_j}{\sigma_{v,j}^2}\right)\\
&-2 n \sum_{i\in N_{gal}}\int \left( \left(\xi_{\delta v}\frac{1}{\bm{\sigma}_v^2}  \xi_{v \delta}\right)_{iV} + \frac{\Delta v_i}{\sigma_{v,i}^2} \left( \xi_{v v} \frac{1} {\bm{\sigma}_v^2} \xi_{v \delta}  \right)_{iV} \right) dV\\
& +n^2  \int  \left(\xi_{\delta v}\frac{1}{\bm{\sigma}_v^2}  \xi_{v \delta}\right)_{V_1V_2}dV_1dV_2.
\end{align}

@@

END NEW STUFF

@@

First, we focus on the second term, $\textbf{x}^T\textbf{C}^{-1}\textbf{x}$. Looking at the definition of $\textbf{C}$ in equation (1), we want to factor $\textbf{C}_0$ into an identity matrix. To do so, we introduce matrix $\textbf{M}$ and $\textbf{X}$. We let $\textbf{M}=\sqrt{\textbf{C}_0}$, and $\textbf{S} = \textbf{M} \textbf{X} \textbf{M}$. Then $\textbf{X}=\textbf{M}^{-1}\textbf{S}\textbf{M}^{-1}$. Now, our matrix $\textbf{C}$ can be re-written in terms of $\textbf{X}$ and $\textbf{M}$:

@@

\textbf{C}^{-1} = \textbf{M}^{-1}(1-A\textbf{X}) \textbf{M}^{-1} = -A \textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1} + \text{const}.

@@

&= -A\left(\textbf{x}^T \textbf{C}_0^{-1}\right)\textbf{S}\left(\textbf{C}_0^{-1}\textbf{x}\right).

@@

-\left(\textbf{x}^T \textbf{C}_0^{-1}\right)\textbf{S}\left(\textbf{C}_0^{-1}\textbf{x}\right) = s_1 +s_2+\hdots +s_6

@@

&=\textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1}

@@

I first sight I wanted to move $\textbf{S}$ across the $\textbf{C}_0^{-1}$ by using a commutator, thinking that it might simplify things a lot. But it turned out that the commutator did not lead to the expected cancelation, so let's do it the hard way. We first do $\textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1}\textbf{S}$, henceforth denoting the inverse of $\bm{\sigma}^2_v\textbf{I}$ as $\bm{\sigma}^{-1}$:

@@

\textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1}\textbf{S} &=

@@

Next, we multiply $\textbf{C}_0^{-1}$ from the right and get

@@

\textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1} &=

@@

\textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1}=

@@

Having obtained the 4 block matrices, we now proceed with the calculation of $\textbf{x}^T\textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1}\textbf{x}$. Similar to what we did for the first order, we break the sum into the following terms (again, summation is within the range of $N_{gal}$ by default unless otherwise stated):

@@

&= -\frac{1}{2}\tr(\textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1}\textbf{S}) \\
